---
title: "Analysis of Montreal Gang Activity"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(igraph)
library(devtools)
library(isnar)
library(dyads)
library(ergm)
```

```{r}
nett.nodes = read.csv("MONTREALGANG.csv")
nett.attr = read.csv("sorted_attrs.csv")

#get rid of title
nett.nodes = nett.nodes[,2:36]

#change NA data to 0 to help with coloring later
nett.attr$Allegiances[which(nett.attr$Allegiances == 99)] = 0
nett.attr$Territories[which(nett.attr$Territories == 99)] = 0
nett.attr$Ethnicity[which(nett.attr$Ethnicity == 99)] = 0


#nodes for network of only friendly ties
friendly_nodes = read.csv("round4/Friendly-Friendly  1.csv")
friendly_nodes = friendly_nodes[2:36]
friendly_nodes = friendly_nodes[1:35,]

#nodes for network of only antagonistic ties
angry_nodes = read.csv("round4/mad-Friendly  1.csv")
angry_nodes = angry_nodes[2:36]
angry_nodes = angry_nodes[1:35,]
```


```{r}
#make full graph

nett.nodes = as.matrix(nett.nodes)
alll = graph_from_adjacency_matrix(nett.nodes)

V(alll)$name = nett.attr[,1]
V(alll)$territory = nett.attr$Territories
V(alll)$ethnicity = nett.attr$Ethnicity
V(alll)$Allegiances = nett.attr$Allegiances
```

```{r} 
#create igraph for positive nodes
friendly_nodes = as.matrix(friendly_nodes)
positive = graph_from_adjacency_matrix(friendly_nodes)

V(positive)$name = nett.attr[,1]
V(positive)$territory = nett.attr$Territories
V(positive)$ethnicity = nett.attr$Ethnicity
V(positive)$Allegiances = nett.attr$Allegiances
```

```{r}
#create igraph for negitive nodes
angry_nodes = as.matrix(angry_nodes)
negative = graph_from_adjacency_matrix(angry_nodes)

V(negative)$name = nett.attr[,1]
V(negative)$territory = nett.attr$Territories
V(negative)$ethnicity = nett.attr$Ethnicity
V(negative)$Allegiances = nett.attr$Allegiances
```

```{r}
#trying different plots

cols = c("dark green", "blue", "red", "black", "yellow", "purple")

#negative by allegiance
plot(negative, vertex.label = NA, vertex.size = 10, 
     edge.arrow.size = 0.1, main = "Graph of Negative by Allegiance", vertex.color = cols[nett.attr$Allegiances+1],layout = layout_with_kk)

legend(x=-1.1, y=-1.1, c("None","Crips", "Bloods", "Other"), pch=21,
       col="#777777", pt.bg=cols, pt.cex=2.5, bty="n", ncol=1)

#positive by allegiance
plot(positive, vertex.label = NA, vertex.size = 10, 
     edge.arrow.size = 0.1, main = "Graph of Positive by Allegiance", vertex.color = cols[nett.attr$Allegiances+1],layout = layout_with_kk)

legend(x=-1.1, y=-1.1, c("None","Crips", "Bloods", "Other"), pch=21,
       col="#777777", pt.bg=cols, pt.cex=2.5, bty="n", ncol=1)
#Full network by ethnicity
plot(alll, vertex.label = NA, vertex.size = 10, 
     edge.arrow.size = 0.1, main = "Graph of All by Ethnicity", vertex.color = cols[nett.attr$Ethnicity+1],layout = layout_with_kk)

legend(x=-1.1, y=-1.1, c("None","Hispanic", "Afro-Candian", "Caucasion", "Asian", "Mixed"), pch=21,
       col="#777777", pt.bg=cols, pt.cex=2.5, bty="n", ncol=1)


```
```{r}
#table with various descriptive statistics
degs = c(mean(degree(alll))/2, mean(degree(positive))/2, mean(degree(negative))/2)
trans = c(transitivity(alll), transitivity(positive), transitivity(negative))
centrality = c(mean(betweenness(alll)), mean(betweenness(positive)), mean(betweenness(negative)))
table1 = matrix(c(degs,trans, (centrality)), nrow = 3, byrow = T)

row.names(table1) = c("Average Degree", "Transitivity", "Centrality")
colnames(table1) = c("Full_Network", "Positive_Network", "Negative_Network")
table1
```


```{r}
#more descriptive statistics
affi = c(ei(alll, "Allegiances"), ei(positive, "Allegiances"),ei(negative, "Allegiances"))
ethni = c(ei(alll, "ethnicity"), ei(positive, "ethnicity"),ei(negative, "ethnicity"))
terr = c(ei(alll, "territory"), ei(positive, "territory"),ei(negative, "territory"))

table2 = matrix(c(affi, ethni, terr), nrow = 3, byrow = T)

row.names(table2) = c("Allegiance", "Ethnicity", "Territory")
colnames(table2) = c("Full_Network", "Positive_Network", "Negative_Network")
table2
```

```{r}
#to run p2 model, need graph as network object, not igraph object
pos2 = as.network(friendly_nodes)
pos2%v%"Ethnicity" = nett.attr$Ethnicity
pos2%v%"Alligiance" = nett.attr$Allegiances
pos2%v%"Territory" = nett.attr$Territories
pos2%v%"Name" = nett.attr[,1]

```

```{r}
#same for negative
neg2 = as.network(angry_nodes)
neg2%v%"Ethnicity" = nett.attr$Ethnicity
neg2%v%"Alligiance" = nett.attr$Allegiances
neg2%v%"Territory" = nett.attr$Territories
neg2%v%"Name" = nett.attr[,1]


```

```{r}
#affiliation positive
Y <- as.matrix(pos2)
group <- get.vertex.attribute(pos2, "Alligiance")
sameGroup <- outer(group, group, "==") * 1

res0 <- p2(Y, sender = NULL, receiver = NULL,density = ~ sameGroup,reciprocity= ~ sameGroup)

tab <- cbind(res[,’Estimate’], res[,’SE’],round(res[,’Estimate’] / res[,’SE’], 3))

colnames(tab) <- c(’Estimate’,’SE’, ’z-stat’)

tab
```

```{r}
#affiliation negative

Y <- as.matrix(neg2)
group2 <- get.vertex.attribute(neg2, "Alligiance")
sameGroup2 <- outer(group, group, "==") * 1

res2 <- p2(Y, sender = NULL, receiver = NULL,density = ~ sameGroup2,reciprocity= ~ sameGroup2)

tab <- cbind(res2[,’Estimate’], res2[,’SE’],round(res2[,’Estimate’] / res[,’SE’], 3))

colnames(tab) <- c(’Estimate’,’SE’, ’z-stat’)

tab


```

```{r}
#needed a clean screenshot for the paper lol
print("       Positive Alligiance")
res0
print("       Negative Alligiance")
res2

```

Look at same group underneath the density. Negative ties are significantly less likely between groups of the same alligiance. Alligiance interestingly has no effect on positive ties.

```{r}
#negative by allegiance and ethnicity
Y <- as.matrix(neg2)
group <- get.vertex.attribute(neg2, "Alligiance")
group2 = get.vertex.attribute(neg2, "Ethnicity")
sameGroup <- outer(group, group, "==") * 1
sameGroup2 =  outer(group2, group2, "==") * 1

res <- p2(Y, sender = NULL, receiver = NULL,density = ~ sameGroup+ sameGroup2,reciprocity= ~ sameGroup+ sameGroup2)

tab <- cbind(res[,’Estimate’], res[,’SE’],round(res[,’Estimate’] / res[,’SE’], 3))

colnames(tab) <- c(’Estimate’,’SE’, ’z-stat’)

tab

```
For negitive, after accounting for gang alligance, ethhnicity is significant, barely. Gangs of the same ethnicity are more likely to have a negative relationship with each other after accounting for alligiance. 

```{r}
#negative by allegiance and territory
Y <- as.matrix(neg2)
group <- get.vertex.attribute(neg2, "Alligiance")
group2 = get.vertex.attribute(neg2, "Territory")
sameGroup <- outer(group, group, "==") * 1
sameGroup2 =  outer(group2, group2, "==") * 1

res3 <- p2(Y, sender = NULL, receiver = NULL,density = ~ sameGroup+ sameGroup2,reciprocity= ~ sameGroup+ sameGroup2)

tab <- cbind(res[,’Estimate’], res[,’SE’],round(res[,’Estimate’] / res[,’SE’], 3))

colnames(tab) <- c(’Estimate’,’SE’, ’z-stat’)

tab
```

```{r}

print("Negative Alligiance and Territory")
res3
```
Gangs in the same territory, fater accounting for gang alligance, are significantly less likely to have negative ties with each other.


```{r}
#trying to see if being part of different allegiances caused more negative ties. 
Y <- as.matrix(neg2)
group2 <- get.vertex.attribute(neg2, "Alligiance")
difGroup2 <- outer(group, group, "!=") * 1

res5 <- p2(Y, sender = NULL, receiver = NULL,density = ~ difGroup2,reciprocity= ~ difGroup2)

tab <- cbind(res2[,’Estimate’], res2[,’SE’],round(res2[,’Estimate’] / res[,’SE’], 3))

colnames(tab) <- c(’Estimate’,’SE’, ’z-stat’)

tab
```

```{r}
print("Negative Alligiance and Ethnicity")
res

```
